Efficient ancestry and mutation simulation with msprime 1.0

### 2021-08-17 paper.tex
Updates from Overleaf

The coalescent
process~\citep{kingman1982coalescent,hudson1983testing,tajima1983evolutionary}
models the ancestry of a set of sampled genomes,
providing a mathematical description of the
genealogical tree that relates the samples to one another.
It has proved to be a powerful model,
and is now central to
population genetics~\citep{hudson1990gene,hein2004gene,wakely2008coalescent}.
The coalescent is a very efficient framework for population genetic
simulation, because it allows us to simulate the genetic ancestry for
a sample from an idealised population model, without explicitly representing
the population in memory or stepping through the generations.
Indeed, \cite{hudson1983testing} independently derived the
% Anyone disagree here? That's the impression I get from the paper.
coalescent \emph{in order to} efficiently simulate data,
and used these simulations to characterise an analytically
intractable distribution.
This inherent efficiency, and the great utility of simulations for a wide
range of purposes, has led to dozens of different tools
being developed over the decades~\citep{carvajal2008simulation,liu2008survey,
arenas2012simulation,yuan2012overview,hoban2012computer,yang2014critical,
peng2015genetic}.

@@

Part of the reason for this poor record of software reuse and modularity is the
lack of standardised file formats, and in particular, the absence of common
library infrastructure to abstract the details of interchanging simulation
data. Although \msprime\ also supports simulating both ancestry and mutations,
the two aspects are functionally independent within the software.
Both ancestry and mutation simulators are present in \msprime\
for reasons of convenience and history,
and could be split into separate packages.
The mature and efficient C and Python interfaces for
\tskit\ make it straightforward to add further information to
an existing file, and because of its efficient data interchange mechanisms,
there is no performance penalty for additional operations
in a different software package.
Thanks to this interoperability, \msprime's
mutation generator can work with \emph{any} \tskit\ tree sequence,
be it simulated using \SLiM~\citep{haller2019slim} or
\fwdpy~\citep{thornton2014cpp}, or estimated from real
data~\citep{kelleher2019inferring,speidel2019method,wohns2021unified}.
It is a modular
component intended to fit into a larger software ecosystem, and
is in no way dependent on \msprime's ancestry simulator.

@@

Simulating mutations in \msprime\ is efficient.
Fig.~\ref{fig-mutations-perf} shows the time required to generate
mutations (using the default JC69 model) on
simulated tree sequences for a variety of mutation
rates as we vary the number of samples
(Fig.~\ref{fig-mutations-perf}A) and the sequence
length (Fig.~\ref{fig-mutations-perf}B).
For example, the longest running simulation in
Fig.~\ref{fig-mutations-perf}B required less than 2 seconds to
generate an average of 1.5 million mutations over 137,081 trees
in a tree sequence with 508,125 edges.
This efficiency for large numbers of trees is possible because
the tree sequence encoding allows us to generate mutations
on an edge-by-edge basis
(see Fig.~\ref{fig-ts-example} and the~\nameref{app-mutation-algorithm}
appendix),
rather than tree-by-tree and branch-by-branch as would otherwise be required.
In the above example from Fig.~\ref{fig-mutations-perf}B,
if we generated mutations tree-by-tree, we would have to iterate over 273,887,838 branches
(since there are 137,081 trees and 1,998 branches in each
tree) rather than 508,125 edges, resulting in $\sim$500 times more work.
% python3 evaluation/generate_mutations_perf_data.py benchmark-single-tree
Even if we have a tree sequence consisting of a single tree
(negating the advantage of working edge-by-edge),
\msprime's mutation generator is still very efficient.
For example, we simulated mutations under the BLOSUM62 amino
acid model for a tree with $10^6$ leaves over $10^4$ sites (resulting
in $\sim$260,000 mutations) in about $0.8$ seconds, including
the time required for file input and output.
We do not attempt a systematic benchmarking of \msprime's
mutation generation code against other methods, because at this scale it is
difficult to disentangle the effects of inefficient input and
output formats from the mutation generation algorithms.
Given these timings, it seems quite unlikely
that generating mutations with \msprime\ would be a bottleneck in any
realistic analysis.

@@

Crossover recombination is implemented in \msprime\ using Hudson's algorithm, which
works backwards in time, generating common ancestor and recombination
events and tracking their effects on segments of ancestral
material inherited from the
sample~\citep{hudson1983properties,hudson1990gene,kelleher2016efficient}.
Common ancestor events merge the ancestral material of two lineages,
and result in coalescences in the marginal trees when ancestral
segments overlap.
Recombination events split the ancestral material for some lineage
at a breakpoint, creating two independent
lineages. Using the appropriate data structures~\citep{kelleher2016efficient},
this process is much more efficient to simulate
than the equivalent left-to-right
approach~\citep{wiuf1999recombination,wiuf1999ancestry}.
In \msprime\ 1.0 recombination rates can vary along a chromosome, allowing
us to simulate recombination hotspots and patterns
of recombination from empirical maps.
The implementation of recombination in \msprime\ is extensively validated
against analytical results~\citep{hudson1983properties,kaplan1985use}
and simulations by \ms, \msHOT\ and \SLiM.

@@

Fig.~\ref{fig-ancestry-perf} is a useful yardstick, allowing us
to predict how long simulations should take for a wide range of
species. Taking a typical chromosome to be 1 Morgan in length,
these plots show, roughly, that simulating chromosome-length
samples from a population of thousands of individuals takes seconds,
while samples from a population of tens of thousands takes minutes.
Simulating whole chromosomes for many species is very fast,
with 1000 samples of chromosome 1 for
\emph{Arabidopsis thaliana} taking less than a second, and a few
minutes for dogs and humans. However, the dependence on $\rho$
\emph{is} quadratic, and if $\rho$ is sufficiently large simulations
may not be feasible. For example,
the \emph{Drosophila melanogaster} chromosome 2L
is about 23.5Mb long with an average recombination rate of around $2.4 \times 10^{-8}$,
so $L \approx 0.57$,
and with $N_e = 1.7 \times 10^6$ \citep{li2006inferring},
$N_e L \approx 10^6$, so extrapolating the curve in Fig.~\ref{fig-ancestry-perf}B
predicts that simulation would require around 177 hours for 1000 samples.
For such large values of $\rho$  we recommend
users consider approximate simulations. Since \msprime\ does not currently
have efficient implementations of approximate coalescent with recombination
models, we recommend using SMC based methods such as \scrm,
particularly for small sample sizes.
In practice, to predict the running time of a given
simulation in \msprime, we recommend that users
measure run time in a series of simulations with short genome lengths and
the desired sample size,
and then predict run time by fitting a quadratic curve to genome length
as in Fig.~\ref{fig-ancestry-perf}.
It is important to note that the quadratic curves in the two
panels of Fig.~\ref{fig-ancestry-perf} are different, and
predicting the run times of days-long simulations
using the timing of seconds-long runs is unlikely to be very accurate.

What about simulations with changing population size?
To understand how run time depends on demography
it helps to consider why run time is quadratic in $\rho$.
At any point in time, \msprime\ must keep track of some number of lineages,
each of which contains some number of chunks of genetic material.
Common ancestor events reduce the number of lineages,
and recombination events increase their number.
However, with long genomes
only a small fraction of the common ancestor events
will involve overlapping segments of ancestry,
and lead to coalescence in the marginal trees.
Such disjoint segments are often far apart (on average, about distance $L/2$),
and so recombine apart again immediately;
it is these large numbers of rapid and inconsequential events that
lead to the quadratic run time.
The maximum number of lineages occurs when
the increase and decrease in numbers of lineages due to
common ancestor and recombination events balance out.
To get an idea of run time we can estimate when this balance occurs.
Suppose that the maximum number of lineages is $M$;
at this time the rate of common ancestor events is $M (M-1) / (4 N_e)$
and the total rate of recombination is $M \ell$,
where $\ell$ is the mean length of genome carried by each lineage
(including ``trapped'' non-ancestral material).
At the maximum, coalescence and recombination rates are equal,
so a typical segment of ancestry will spend roughly half its time
in a lineage with at least one other such segment---and,
since such lineages carry at least two segments,
at most one-third of the lineages carry long trapped segments of ancestry.
Since the maximum number of lineages is reached
very quickly~\citep{nelson2020accounting},
this implies that $\ell \approx L / 6$.
Setting the rates of recombination and common ancestor events
to be equal and solving for $M$,
we find that $M$ is roughly equal to $L N_e$.
The number of lineages then decreases gradually from this maximum
on the coalescent time scale, and therefore over roughly $2 N_e$ generations.
Since the total rate of events when the maximum number of lineages is
present is roughly $L^2 N_e / 6$,
then the total number of events is proportional to $(L N_e)^2$---i.e.,
proportional to $\rho^2$.

@@

Fig.~\ref{fig-dtwf-perf} shows that \msprime\ simulates the DTWF
more quickly and requires substantially less memory than
\ARGON~\citep{palamara2016argon}, a specialised DTWF simulator.
However, the generation-by-generation approach of the DTWF is less
efficient than the coalescent with recombination when the
number of lineages is significantly less than the population size
(the regime where the coalescent is an accurate approximation),
which usually happens in the quite recent
past~\cite{bhaskar2014distortion}.
We therefore support changing the simulation model during a simulation
so that we can run hybrid simulations, as proposed by~\cite{bhaskar2014distortion}.
Any number of different simulation models can be combined, allowing for the
flexible choice of simulation scenarios.
As the discrete time Wright-Fisher model improves accuracy of genealogical
patterns in the recent past, we can simulate the recent history using this
model and then switch to the standard coalescent to more efficiently simulate
the more ancient history.

@@

More visibly to the end user, we also have a high standard for documentation,
with precise, comprehensive, and cross-linked documentation
that is automatically built from the code base
and served through the website \texttt{https://tskit.dev}.
With the goal of lowering the entry barrier to new users,
we have invested significant effort in writing examples and introductions,
and making common tasks discoverable.
We also view contributions to documentation as equally important to the project
as writing code or designing methods---what use would it be
to write reliable, stable software if no-one used it?

@@

First, mutations are placed on the tree sequence under an inhomogeneous Poisson model
by applying them independently to each edge.
If an edge spans a region $[a, b)$ of the genome
and connected parent and child nodes with times $s < t$,
and the mutation rate locally is $\mu$,
then the number of mutations on the edge is Poisson with mean $\mu (t-s) (b-a)$,
and each mutation is placed independently at a position chosen uniformly in $[a,b)$
and a time uniformly in $[s,t)$.
In a discrete genome, all positions are integers and so more than one mutation
may occur at the same position on the same edge.
Otherwise (i.e., for an infinite-sites model),
positions are rejection sampled to obtain a unique floating-point number.
If an edge spans a region of the genome with more than one mutation rate,
this is done separately for each distinct region.
% I am a bit confused with the following lines - as the section in the main paper says, algo is done on edges,
% but this is skipped here and the way its written it makes it confusing - #edges are propto #marginal_trees
% so there should be no difference (the way I read this part), but in fact the difference on working with edges
% or each tree and its branches can be large
The algorithm scales with the number of edges in the tree sequence.
Since the number of edges is proportional to the number of marginal trees,
an algorithm that applies mutations separately to each marginal tree
would take at least as many steps,
and would require additional bookkeeping necessary to place mutations uniformly on a tree.

Next, alleles are chosen for each mutation.
If the site was not previously mutated, then a new ancestral allele is chosen for the site,
according to an input distribution of ancestral state allele probabilities.
Then, each mutation on the tree is considered in turn,
% I don't understand what is meant by "older mutations are considered first"?
% Aren't past/older mutations irrelevant given the current state (=parental allele) and a transition matrix?
in order so that older mutations are considered first,
and a derived allele is chosen based on the parental allele
(which may be the ancestral allele or the derived allele of a previous mutation)
% What is meant by "some randomness"?
and some randomness.
Finally, information about the mutations are recorded in the site and mutation tables
of the tree sequence.

@@

% which algo do you refer above? InfiniteAlleles or the general version above the above paragraph? Better to spell it out.
Since the algorithm above applies mutations at a single rate
independent of ancestral state,
a model in which different alleles mutate at different rates
must necessarily produce some silent mutations,
i.e., mutations in which the derived allele is equal to the parental allele.
To illustrate this, consider a mutation model on the binary alphabet
$\{X, Y\}$ in which $X$ mutates to $Y$ at rate $\alpha$
and $Y$ mutates to $X$ at rate $\beta$, with $\beta < \alpha$.
To implement this, first place mutations at the largest total rate, which is $\alpha$.
Then, at each site, choose an ancestral allele from the ``root distribution''
(a natural choice would be to choose from the stationary distribution,
$X$ with probability $\beta / (\alpha + \beta)$),
and for each mutation, choose a derived allele as follows:
if the parental allele is $X$, then choose $Y$;
if the parental allele is $Y$, then choose $X$ with probability $\beta/(\alpha + \beta)$,
and choose $Y$ otherwise.
This produces the correct distribution by Poisson thinning:
a Poisson process with rate $\alpha$ in which each point is discarded independently
with probability $\beta / (\alpha + \beta)$ is equivalent to a Poisson
process with rate $\beta$.
All finite-state models (implemented under the generic \texttt{MatrixMutationModel} class)
work in this way: mutations are placed at the maximum mutation rate,
% I think I understand what you want to say here, but it would be clearer if an example would have been given - similar to the next section.
and then some silent mutations will result.

>>>>

The coalescent
process~\citep{kingman1982coalescent,hudson1983testing,tajima1983evolutionary}
models the ancestry of a set of sampled genomes,
providing a mathematical description of the
genealogical tree that relates the samples to one another.
It has proved to be a powerful model,
and is now central to
population genetics~\citep{hudson1990gene,hein2004gene,wakely2008coalescent}.
The coalescent is an efficient framework for population genetic
simulation, because it allows us to simulate the genetic ancestry for
a sample from an idealised population model, without explicitly representing
the population in memory or stepping through the generations.
Indeed, \cite{hudson1983testing} independently derived the
% Anyone disagree here? That's the impression I get from the paper.
coalescent \emph{in order to} efficiently simulate data,
and used these simulations to characterise an analytically
intractable distribution.
This inherent efficiency, and the great utility of simulations for a wide
range of purposes, has led to dozens of different tools
being developed over the decades~\citep{carvajal2008simulation,liu2008survey,
arenas2012simulation,yuan2012overview,hoban2012computer,yang2014critical,
peng2015genetic}.

@@

Part of the reason for this poor record of software reuse and modularity is the
lack of standardised file formats, and in particular, the absence of common
library infrastructure to abstract the details of interchanging simulation
data. Although \msprime\ also supports simulating both ancestry and mutations,
the two aspects are functionally independent within the software; both ancestry and mutation simulators are present in \msprime\
for reasons of convenience and history,
and could be split into separate packages.
The mature and efficient C and Python interfaces for
\tskit\ make it straightforward to add further information to
an existing file, and because of its efficient data interchange mechanisms,
there is no performance penalty for additional operations
in a different software package.
Thanks to this interoperability, \msprime's
mutation generator can work with \emph{any} \tskit\ tree sequence,
be it simulated using \SLiM~\citep{haller2019slim} or
\fwdpy~\citep{thornton2014cpp}, or estimated from real
data~\citep{kelleher2019inferring,speidel2019method,wohns2021unified}.
It is a modular
component intended to fit into a larger software ecosystem, and
is in no way dependent on \msprime's ancestry simulator.

@@

Simulating mutations in \msprime\ is efficient.
Fig.~\ref{fig-mutations-perf} shows the time required to generate
mutations (using the default JC69 model) on
simulated tree sequences for a variety of mutation
rates as we vary the number of samples
(Fig.~\ref{fig-mutations-perf}A) and the sequence
length (Fig.~\ref{fig-mutations-perf}B).
For example, the longest running simulation in
Fig.~\ref{fig-mutations-perf}B required less than 2 seconds to
generate an average of 1.5 million mutations over 137,081 trees
in a tree sequence with 508,125 edges.
This efficiency for large numbers of trees is possible because
the tree sequence encoding allows us to generate mutations
on an edge-by-edge basis
(see Fig.~\ref{fig-ts-example} and the~\nameref{app-mutation-algorithm}
appendix),
rather than tree-by-tree and branch-by-branch as would otherwise be required.
In the above example from Fig.~\ref{fig-mutations-perf}B,
if we generated mutations tree-by-tree, we would have to iterate over 273,887,838 branches
(since there are 137,081 trees and 1,998 branches in each
tree) rather than 508,125 edges, resulting in $\sim$500 times more work.
% python3 evaluation/generate_mutations_perf_data.py benchmark-single-tree
Even if we have a tree sequence consisting of a single tree
(negating the advantage of working edge-by-edge),
\msprime's mutation generator is still very efficient.
For example, we simulated mutations under the BLOSUM62 amino
acid model for a tree with $10^6$ leaves over $10^4$ sites (resulting
in $\sim$260,000 mutations) in about $0.8$ seconds, including
the time required for file input and output.
We do not attempt a systematic benchmarking of \msprime's
mutation generation code against other methods, because at this scale it is
difficult to disentangle the effects of inefficient input and
output formats from the mutation generation algorithms.
Given these timings, it seems unlikely
that generating mutations with \msprime\ would be a bottleneck in any
realistic analysis.

@@

Crossover recombination is implemented in \msprime\ using Hudson's algorithm, which
works backwards in time, generating common ancestor and recombination
events and tracking their effects on segments of ancestral
material inherited from the
sample~\citep{hudson1983properties,hudson1990gene,kelleher2016efficient}.
Common ancestor events merge the ancestral material of two lineages,
and result in coalescences in the marginal trees when ancestral
segments overlap.
Recombination events split the ancestral material for some lineage
at a breakpoint, creating two independent
lineages. Using the appropriate data structures~\citep{kelleher2016efficient},
this process is much more efficient to simulate
than the equivalent left-to-right
approach~\citep{wiuf1999recombination,wiuf1999ancestry}.
In \msprime\ 1.0, recombination rates can vary along a chromosome, allowing
us to simulate recombination hotspots and patterns
of recombination from empirical maps.
The implementation of recombination in \msprime\ is extensively validated
against analytical results~\citep{hudson1983properties,kaplan1985use}
and simulations by \ms, \msHOT\ and \SLiM.

@@

Fig.~\ref{fig-ancestry-perf} is a useful yardstick, allowing us
to predict how long simulations should take for a wide range of
species. Taking a typical chromosome to be 1 Morgan in length,
these plots show, roughly, that simulating chromosome-length
samples from a population of thousands of individuals takes seconds,
while samples from a population of tens of thousands takes minutes.
Simulating whole chromosomes for many species is very fast,
with 1000 samples of chromosome 1 for
\emph{Arabidopsis thaliana} taking less than a second, and a few
minutes for dogs and humans. However, the dependence on $\rho$
\emph{is} quadratic, and if $\rho$ is sufficiently large simulations
may not be feasible. For example,
the \emph{Drosophila melanogaster} chromosome 2L
is about 23.5Mb long with an average recombination rate of around $2.4 \times 10^{-8}$,
so $L \approx 0.57$,
and with $N_e = 1.7 \times 10^6$ \citep{li2006inferring},
$N_e L \approx 10^6$, so extrapolating the curve in Fig.~\ref{fig-ancestry-perf}B
predicts that simulation would require around 177 hours for 1000 samples.
For such large values of $\rho$  we recommend
users consider approximate simulations. Since \msprime\ does not currently
have efficient implementations of approximate coalescent with recombination
models, in these cases we recommend using SMC based methods such as \scrm,
particularly if sample sizes are small.
In practice, to predict the running time of a given
simulation in \msprime, we recommend that users
measure run time in a series of simulations with short genome lengths and
the desired sample size,
and then predict run time by fitting a quadratic curve to genome length
as in Fig.~\ref{fig-ancestry-perf}.
It is important to note that the quadratic curves in the two
panels of Fig.~\ref{fig-ancestry-perf} are different, and
predicting the run times of days-long simulations
using the timing of seconds-long runs is unlikely to be very accurate.

What about simulations with changing population size?
To understand how run time depends on demography
it helps to consider why run time is quadratic in $\rho$.
At any point in time, \msprime\ must keep track of some number of lineages,
each of which contains some number of chunks of genetic material.
Common ancestor events reduce the number of lineages,
and recombination events increase their number.
However, with long genomes,
only a small fraction of the common ancestor events
will involve overlapping segments of ancestry
and lead to coalescence in the marginal trees.
Such disjoint segments are often far apart (on average, about distance $L/2$),
and so recombine apart again immediately;
it is these large numbers of rapid and inconsequential events that
lead to the quadratic run time.
The maximum number of lineages occurs when
the increase and decrease in numbers of lineages due to
common ancestor and recombination events balance out.
To get an idea of run time we can estimate when this balance occurs.
Suppose that the maximum number of lineages is $M$;
at this time the rate of common ancestor events is $M (M-1) / (4 N_e)$
and the total rate of recombination is $M \ell$,
where $\ell$ is the mean length of genome carried by each lineage
(including ``trapped'' non-ancestral material).
At the maximum, coalescence and recombination rates are equal,
so a typical segment of ancestry will spend roughly half its time
in a lineage with at least one other such segment---and,
since such lineages carry at least two segments,
at most one-third of the lineages carry long trapped segments of ancestry.
Since the maximum number of lineages is reached
very quickly~\citep{nelson2020accounting},
this implies that $\ell \approx L / 6$.
Setting the rates of recombination and common ancestor events
to be equal and solving for $M$,
we find that $M$ is roughly equal to $L N_e$.
The number of lineages then decreases gradually from this maximum
on the coalescent time scale, and therefore over roughly $2 N_e$ generations.
Since the total rate of events when the maximum number of lineages is
present is roughly $L^2 N_e / 6$,
then the total number of events is proportional to $(L N_e)^2$---i.e.,
proportional to $\rho^2$.

@@

Fig.~\ref{fig-dtwf-perf} shows that \msprime\ simulates the DTWF
more quickly and requires substantially less memory than
\ARGON~\citep{palamara2016argon}, a specialised DTWF simulator.
However, the generation-by-generation approach of the DTWF is less
efficient than the coalescent with recombination when the
number of lineages is significantly less than the population size
(the regime where the coalescent is an accurate approximation),
which usually happens in the quite recent
past~\citep{bhaskar2014distortion}.
We therefore support changing the simulation model during a simulation
so that we can run hybrid simulations, as proposed by~\cite{bhaskar2014distortion}.
Any number of different simulation models can be combined, allowing for the
flexible choice of simulation scenarios.
As the discrete time Wright-Fisher model improves accuracy of genealogical
patterns in the recent past, we can simulate the recent history using this
model and then switch to the standard coalescent to more efficiently simulate
the more ancient history.

@@

More visibly to the end user, we also have a high standard for documentation,
with precise, comprehensive, and cross-linked documentation
that is automatically built from the code base
and served through the website \texttt{https://tskit.dev}.
With the goal of lowering the entry barrier to new users,
we have invested significant effort in writing examples and introductions,
and making common tasks discoverable.
We also view contributions to documentation as equally important to the project
as writing code or designing methods: what use would it be
to write reliable, stable software if no-one used it?

@@

First, mutations are placed on the tree sequence under an inhomogeneous Poisson model
by applying them independently to each edge.
If an edge spans a region $[a, b)$ of the genome
and connected parent and child nodes with times $s < t$,
and the mutation rate locally is $\mu$,
then the number of mutations on the edge is Poisson with mean $\mu (t-s) (b-a)$,
and each mutation is placed independently at a position chosen uniformly in $[a,b)$
and a time uniformly in $[s,t)$.
In a discrete genome, all positions are integers and so more than one mutation
may occur at the same position on the same edge.
Otherwise (i.e., for an infinite-sites model),
positions are rejection sampled to obtain a unique floating-point number.
If an edge spans a region of the genome with more than one mutation rate,
this is done separately for each distinct region.
The algorithm scales with the number of edges in the tree sequence.
Since the number of edges is proportional to the number of marginal trees,
an algorithm that applies mutations separately to each marginal tree
would take at least as many steps,
and would require additional bookkeeping necessary to place mutations uniformly on a tree.

Next, alleles are chosen for each mutation.
If the site was not previously mutated, then a new ancestral allele is chosen for the site,
according to an input distribution of ancestral state allele probabilities.
Then, each mutation on the tree is considered in turn,
in order so that older mutations are considered first,
and a derived allele is chosen based on the parental allele
(which may be the ancestral allele or the derived allele of a previous mutation)
and some randomness.
Finally, information about the mutations are recorded in the site and mutation tables
of the tree sequence.

@@

Since the algorithm above applies mutations at a single rate
independent of ancestral state,
a model in which different alleles mutate at different rates
must necessarily produce some silent mutations,
i.e., mutations in which the derived allele is equal to the parental allele.
To illustrate this, consider a mutation model on the binary alphabet
$\{X, Y\}$ in which $X$ mutates to $Y$ at rate $\alpha$
and $Y$ mutates to $X$ at rate $\beta$, with $\beta < \alpha$.
To implement this, first place mutations at the largest total rate, which is $\alpha$.
Then, at each site, choose an ancestral allele from the ``root distribution''
(a natural choice would be to choose from the stationary distribution,
$X$ with probability $\beta / (\alpha + \beta)$),
and for each mutation, choose a derived allele as follows:
if the parental allele is $X$, then choose $Y$;
if the parental allele is $Y$, then choose $X$ with probability $\beta/(\alpha + \beta)$,
and choose $Y$ otherwise.
This produces the correct distribution by Poisson thinning:
a Poisson process with rate $\alpha$ in which each point is discarded independently
with probability $\beta / (\alpha + \beta)$ is equivalent to a Poisson
process with rate $\beta$.
All finite-state models (implemented under the generic \texttt{MatrixMutationModel} class)
work in this way: mutations are placed at the maximum mutation rate,
and then some silent mutations will result.
